Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-11-24 text/supp.tex
Edits based on Jon's comments

\caption{Summary statistics for the filtered Global Population Dynamics
  Database time series arranged by taxonomic class. Columns are: number of
  populations, number of taxonomic orders, numbers of species, median time
  series length, total number of interpolated time steps, total number of
  substituted zeros, and total number of time steps.}

@@

\caption{All populations with Pr$(\nu < 10) > 0.5$ in the base heavy-tailed
  Gompertz population dynamics model. Shown are the log abundance time series,
  population descriptions, Global Population Dynamics Database Main IDs,
  citation for the data source or separate verification literature, a
  description of the cause of the black swan events (if known), the
  probability of heavy tails as calculated by our model, and median estimate
  of $\nu$ from our model with 90\% quantile credible intervals indicated in
  parentheses. Red dots on the time series indicate downward black-swan events
  and blue values indicate upward black-swan events that have a $10 \cdot
  10^{-4}$ probability or less of occurring if the population dynamics were
  explained by a Gompertz model with normally distributed process noise with a
  standard deviation equal to the scale parameter in the fitted t
  distribution.}\\

@@

\caption{(fishes, crustaceans, gastropods, sharks). All filtered time series
  used in our analysis. The abundances are shown on a log10 vertical axis.
  Throughout this figure, red dots indicate values that were interpolated and
  blue dots indicate values that were recorded as zero but were set to the
  next lowest observed abundance. Numbers before each species name are the
  GPDD Main ID numbers.}

@@

\caption{ Probability density of the Bayesian priors for the Gompertz models.
  From left to right and then top to bottom: (1) per capita growth rate at
  $\log$(abundance) = $0$: $\lambda \sim \mathrm{Normal}(0, 10^2)$; (2) scale
  parameter of t-distribution process noise: $\sigma \sim \mathrm{Half\mhyphen
    Cauchy} (0, 2.5)$; (3) t-distribution degrees of freedom parameter: $\nu
  \sim \mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2)$; (4)
  AR1 correlation coefficient of residuals: $\phi \sim \mathrm{Truncated
    \mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. Not shown
  is $b$, the density dependence parameter: $b \sim \mathrm{Uniform}(-1, 2)$.
  The $\nu$ panel also shows two alternative priors: a weaker prior $\nu \sim
  \mathrm{Truncated\mhyphen Exponential}(0.005, \mathrm{min.} = 2)$, and a
  stronger prior $\nu \sim \mathrm{Truncated\mhyphen Exponential}(0.02,
  \mathrm{min.} = 2)$. The inset panel shows the same data but with a
  log-transformed x axis. Note that the base and weaker priors are relatively
  flat within the region of $\nu < 20$ that we are concerned with. }

@@

\caption{Estimates of $\nu$ from alternative models plotted against the base
  Gompertz model estimates of $\nu$. Shown are medians of the posterior (dots)
  and 50\% credible intervals (segments). The diagonal line indicates a
  one-to-one relationship. Different colours indicate various taxonomic
  classes. The grey-shaded regions indicate regions of disagreement if $\nu =
  10$ is taken as a threshold of heavy-tailed dynamics. The Gompertz
  observation error model assumes a fixed standard deviation of observation
  error of $0.2$ on a log scale.}

@@

\caption{Estimates of $\nu$ from Gompertz models with alternative priors on
  $\nu$. Shown are medians of the posterior (dots) and 50\% credible intervals
  (segments). The diagonal line indicates a one-to-one relationship. Different
  colours indicate various taxonomic classes. The grey-shaded regions indicate
  regions of disagreement if $\nu = 10$ is taken as a threshold of
  heavy-tailed dynamics. The base, weaker, and stronger priors on $\nu$ are
  illustrated in Fig.~\ref{fig:priors}. In general, the estimates are nearly
  identical in cases where the data are informative about low values of $\nu$.
  When the data are less informative about low values of $\nu$, the prior can
  slightly pull the estimates of $\nu$ towards higher or lower values.}

@@

\caption{Testing the ability to estimate $\nu$ (top panels) and the scale
  parameter of the process deviations (bottom panels) for a given number of
  samples (columns) drawn from a distribution with a given true $\nu$ value
  (rows). The red lines indicate the true population value. When a small
  number of samples are drawn there may not be samples sufficiently far into
  the tails to recapture the true $\nu$ value; however, heavy tails are still
  distinguished from normal tails in most cases, even with only 25 or 50
  samples.}

@@

\caption{ Simulation testing the Gompertz estimation model when the process
  deviation draws were chosen so that $\nu$ could be estimated close to the
  true value outside the full population model (``effective $\nu$'' within a
  CV of 0.2 of specified $\nu$). Upper panels show the distribution of median
  $\widehat{\nu}$ across 20 simulation runs. Lower panels show the
  distribution of Pr($\nu < 10$) across 20 simulation runs. We ran the
  simulations across three population (``true'') $\nu$ values (3, 5, and
  $1\cdot 10^9$, i.e.\ approximately normal) and four scenarios: (1) 100 time
  steps and no observation error, (2) 50 time steps and no observation error,
  (3) 50 time steps and observation error drawn from $\mathrm{Normal} (0,
  0.2^2)$ but ignored, and (4) 50 time steps with observation error in which
  the quantity of observation error was assumed known. Within each scenario
  the dots represent stochastic draws from the true population distributions
  combined with model fits. Underlayed boxplots show the median, interquartile
  range, and $1.5$ times the interquartile range. }

>>>>

\caption[Summary statistics for the filtered Global Population Dynamics
  Database time series arranged by taxonomic class.]{Summary statistics for the filtered Global Population Dynamics
  Database time series arranged by taxonomic class. Columns are: number of
  populations, number of taxonomic orders, numbers of species, median time
  series length, total number of interpolated time steps, total number of
  substituted zeros, and total number of time steps.}

@@

\caption[All populations with Pr$(\nu < 10) > 0.5$ in the base heavy-tailed
  Gompertz population dynamics model.]{All populations with Pr$(\nu < 10) > 0.5$ in the base heavy-tailed
  Gompertz population dynamics model. Shown are the log abundance time series,
  population descriptions, Global Population Dynamics Database Main IDs,
  citation for the data source or separate verification literature, a
  description of the cause of the black swan events (if known), the
  probability of heavy tails as calculated by our model, and median estimate
  of $\nu$ from our model with 90\% quantile credible intervals indicated in
  parentheses. Red dots on the time series indicate downward black-swan events
  and blue values indicate upward black-swan events that have a $10 \cdot
  10^{-4}$ probability or less of occurring if the population dynamics were
  explained by a Gompertz model with normally distributed process noise with a
  standard deviation equal to the scale parameter in the fitted t
  distribution.}\\

@@

\caption[All filtered time series used in our analysis.]{(fishes, crustaceans,
  gastropods, sharks). All filtered time series used in our analysis. The
  abundances are shown on a log10 vertical axis. Throughout this figure, red
  dots indicate values that were interpolated and blue dots indicate values
  that were recorded as zero but were set to the next lowest observed
  abundance. Numbers before each species name are the GPDD Main ID numbers.}

@@

\caption[Probability density of the Bayesian priors for the Gompertz models.]{
  Probability density of the Bayesian priors for the Gompertz models. From
  left to right and then top to bottom: (1) per capita growth rate at
  $\log$(abundance) = $0$: $\lambda \sim \mathrm{Normal}(0, 10^2)$; (2) scale
  parameter of t-distribution process noise: $\sigma \sim \mathrm{Half\mhyphen
    Cauchy} (0, 2.5)$; (3) t-distribution degrees of freedom parameter: $\nu
  \sim \mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2)$; (4)
  AR1 correlation coefficient of residuals: $\phi \sim \mathrm{Truncated
    \mhyphen Normal}(0, 1, \mathrm{min.} = -1, \mathrm{max.} = 1)$. Not shown
  is $b$, the density dependence parameter: $b \sim \mathrm{Uniform}(-1, 2)$.
  The $\nu$ panel also shows two alternative priors: a weaker prior $\nu \sim
  \mathrm{Truncated\mhyphen Exponential}(0.005, \mathrm{min.} = 2)$, and a
  stronger prior $\nu \sim \mathrm{Truncated\mhyphen Exponential}(0.02,
  \mathrm{min.} = 2)$. The inset panel shows the same data but with a
  log-transformed x axis. Note that the base and weaker priors are relatively
  flat within the region of $\nu < 20$ that we are concerned with. }

@@

\caption[Estimates of $\nu$ from alternative models plotted against the base
Gompertz model estimates of $\nu$.]{Estimates of $\nu$ from alternative models
  plotted against the base Gompertz model estimates of $\nu$. Shown are
  medians of the posterior (dots) and 50\% credible intervals (segments). The
  diagonal line indicates a one-to-one relationship. Different colours
  indicate various taxonomic classes. The grey-shaded regions indicate regions
  of disagreement if $\nu = 10$ is taken as a threshold of heavy-tailed
  dynamics. The Gompertz observation error model assumes a fixed standard
  deviation of observation error of $0.2$ on a log scale.}

@@

\caption[Estimates of $\nu$ from Gompertz models with alternative priors on
$\nu$.]{Estimates of $\nu$ from Gompertz models with alternative priors on
  $\nu$. Shown are medians of the posterior (dots) and 50\% credible intervals
  (segments). The diagonal line indicates a one-to-one relationship. Different
  colours indicate various taxonomic classes. The grey-shaded regions indicate
  regions of disagreement if $\nu = 10$ is taken as a threshold of
  heavy-tailed dynamics. The base, weaker, and stronger priors on $\nu$ are
  illustrated in Fig.~\ref{fig:priors}. In general, the estimates are nearly
  identical in cases where the data are informative about low values of $\nu$.
  When the data are less informative about low values of $\nu$, the prior can
  slightly pull the estimates of $\nu$ towards higher or lower values.}

@@

\caption[Testing the ability to estimate $\nu$  and the scale parameter of the
process deviations for a given number of samples drawn from a distribution
with a given true $\nu$ value.]{Testing the ability to estimate $\nu$ (top
  panels) and the scale parameter of the process deviations (bottom panels)
  for a given number of samples (columns) drawn from a distribution with a
  given true $\nu$ value (rows). The red lines indicate the true population
  value. When a small number of samples are drawn there may not be samples
  sufficiently far into the tails to recapture the true $\nu$ value; however,
  heavy tails are still distinguished from normal tails in most cases, even
  with only 25 or 50 samples.}

@@

\caption[Simulation testing the Gompertz estimation model when the process
deviation draws were chosen so that $\nu$ could be estimated close to the true
value outside the full population model (``effective $\nu$'' within a CV of
0.2 of specified $\nu$).]{Simulation testing the Gompertz estimation model
  when the process deviation draws were chosen so that $\nu$ could be
  estimated close to the true value outside the full population model
  (``effective $\nu$'' within a CV of 0.2 of specified $\nu$). Upper panels
  show the distribution of median $\widehat{\nu}$ across 20 simulation runs.
  Lower panels show the distribution of Pr($\nu < 10$) across 20 simulation
  runs. We ran the simulations across three population (``true'') $\nu$ values
  (3, 5, and $1\cdot 10^9$, i.e.\ approximately normal) and four scenarios:
  (1) 100 time steps and no observation error, (2) 50 time steps and no
  observation error, (3) 50 time steps and observation error drawn from
  $\mathrm{Normal} (0,
  0.2^2)$ but ignored, and (4) 50 time steps with observation error in which
    the quantity of observation error was assumed known. Within each scenario
    the dots represent stochastic draws from the true population distributions
    combined with model fits. Underlayed boxplots show the median,
    interquartile range, and $1.5$ times the interquartile range. }
